The Symmetry of Partner Modelling

### 2016-02-04 mutual-modeling.tex
Proof-reading sections 1 2 3

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{{\textit{e.g.\ }}}
\newcommand{\al}{{\textit{et al.\ }}}

@@

Collaborative learning has often been associated to the construction of  a
shared   understanding of the situation at hand. The psycholinguistics
mechanisms at work while establishing common grounds are the object of
scientific controversy. We postulate that collaborative tasks require some
level of mutual modelling, i.e. that each partner needs some model of what
the other partners know/want/intend at a given time.  We use the term 'some
model' to stress the fact that this model is not necessarily details or
complete, but that we acquire some representations of the persons we
interact with. The question we address is: Does the quality of the partner
model depend upon the modeler's ability to represent his or her partner,
upon the modelee's ability to make his state clear to the modeler or upon
the quality of their interactions? We address this question by analyzing
differences between the  accuracy of the models built by different team
members. We therefore report 5 experiments on collaborative problem solving
or collaborative learning that vary in terms of tasks (how important it is
to build an accurate model) and settings (how difficult it is to build an
accurate model). In 4 studies, the accuracy of the model that A built about
B was correlated with the accuracy of the model that B built about A,  which
seems to imply that the quality of interactions matters more than individual
abilities when building mutual models.  However, these finding do not
discard the fact individual abilities also  participate in the quality of
modelling process .

@@

From its very beginning, CSCL research has been following the suggestion by
\citet{roschelle1995construction} that collaborative learning has something to
do with the process of constructing and maintaining a \emph{shared
understanding} of the task at hand. Building a shared/mutual understanding
refers to the upper class of collaborative learning situations, those in which
students should build upon each other's understanding to refine their own
understanding: what is expected to produce learning is not the mere fact that
two students build the same understanding but the cognitive effort they have to
engage to build this shared understanding~\citep{schwartz1995emergence}. This
effort can be observed by the frequency of rich interactions, \ie interactions
whose occurrence has been related to learning: (self-) explanations in cognitive
science~\citep{chi1989self, webb1991task}, conflict resolution in
socio-cognitive theories~\citep{doise1975social}, mutual
regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
construction of a shared understanding has been investigated for several decades
in psycholinguistics, under the  notion of
\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned in the learning
sciences. Grounding mechanisms  are appropriate for explains conversational
events, such referential failures in short dialogue episodes, but they do
hardly predict deeper phenomena such as \emph{conceptual change} (\ie the
acquisition, acceptation and integration of a new belief into one's mental
model) over longer sessions~\citep{dillenbourg2006sharing}. The cumulative
effect of grounding episodes can probably be better understood from a
socio-cultural perspective:

@@

Paradoxically, while Clark's theory is somewhat too linguistic from a learning
viewpoint, it is criticized at the same time as being too cognitivist by some
psycholinguists, \ie as overestimating the amount of shared knowledge and mutual
representations actually necessary to conduct a dialogue. The fundamental issue,
as old as philosophy, is the degree of coupling between the different levels of
dialogue, mostly between the lexical/syntactical level and the deeper semantic
levels. \citet{pickering2006alignment} argue that the mutual understanding
starts mostly with a \emph{superficial alignment} at the level of the linguistic
representations, due to priming mechanisms, and that this local alignment may --
in some cases -- lead to a \emph{global alignment} of the semantic level
(\emph{deep grounding}).  For these authors, the convergence in dialogue, and
even the repair of some misunderstandings, is explained by this mimetic behavior
more than by a monitoring of each other's knowledge: \emph{``...interlocutors do
not need to monitor and develop full common ground as a regular, constant
part of routine conversation, as it would be unnecessary and far too costly.
Establishment of full common ground is, we argue, a specialized and
non-automatic process that is used primarily in times of difficulty (when
radical misalignment becomes apparent).''} (p. 179).  This view is actually not
incompatible with Clark's \emph{grounding
criterion}~\citep{clark1989contributing}: the degree of shared understanding
that peers need to reach depends upon the task they perform. For instance, a
dialogue between two surgeons might rely on superficial alignment if they talk
about their friends but has to guarantee accurate common grounds when talking
about which intervention will be conducted in which way on which patient.  In
this paper, we operationalized the grounding criterion, i.e. the necessity for
accurate modelling, as the correlation between the accuracy of partner models
and various measures of team performance.

This interesting cognitive science debate occurred mostly outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes: \textbf{Is linguistic alignment sufficient to trigger conceptual
change?} \textbf{Does negotiation of meaning only occur when partners monitor
and diagnose each other's knowledge?} If the ratio between shallow alignment and
deep grounding depends upon the task, and if deep grounding is a condition for
learning, then the pedagogical challenge is indeed to design tasks that require
deep grounding. Most empirical studies on grounding and alignment are conducted
with simple referencing tasks such as asking the way to the train station
or helping the peer to choose a picture among many. In the studies we report here, we
explore several richer tasks such as arguing on a sensitive issue or building a
concept map.

@@

%\paragraph{Theory of Mind}
%
%Theory of Mind (originally defined in~\citep{premack1978does}) is the cognitive
%ability that a subject possesses to represent the mental state of another
%agent, possibly including knowledge that contradicts the subject's own model: for
%example, a book can be at the same time \emph{visible} for myself, and \emph{not
%visible} for you.
%
%Children develop this skill, which is essential to understand others'
%perspectives during interactions, around the age of three. It supposes the
%ability to build, store and retrieve separate models of the knowledge of the
%interactors.
%
%One classical application of this cognitive skill is the so-called
%\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
%experiment)~\citep{Leslie2000}: a child is asked to watch a scene where two
%people, $A$ and $B$, manipulate objects. Then $A$ leaves and $B$ hides away one
%object. When $A$ comes back, we ask the child ``where do you think $A$ will
%look for the object?''. Before acquiring a theory of mind, children are not
%able to separate their own (true) model of the world (where they know that
%the object was hidden) from the model of A, which contains \emph{false
%beliefs} on the world ($A$ still thinks the object is at its original
%position since he did not see $B$ hiding it).

@@

In summary, the controversy around the cognitive depth of shared understanding
pertains to psycholinguistic, which investigates 'natural conversations, but  in
collaborative learning, the situation is different; tasks are precisely designed
for requiring a deeper negotiation of meaning.  The craft of CSCL is to design
collaborative situations that create interdependence, avoid group think, and
allows students to detect any 'illusion of shared understanding' (Cherubini et
al, 2005). Our question is hence not  anymore"do peers build a shared
understanding ?", but becomes "when peers have to build a shared understanding,
how do they do it ?".  This question addresses the mechanisms of grounding, in
which the basic sequence is: make a proposition, detect misunderstanding and, if
any, repair them.  This paper focuses on the middle part, the detection of
misunderstandings.  Detecting a misunderstanding means that the emitter of a
message identifies a mismatch between his communicative intention and the way
his message is understood by his or her partner. Detecting one peer's
understanding is investigate hereafter under the general umbrella of partner
modelling.

%TODO
Cherubini, M., Van Der Pol, J., \& Dillenbourg, P. (2005). Grounding is not
shared understanding: Distinguishing grounding at an utterance and knowledge
level. In CONTEXT'05, the Fifth International and Interdisciplinary Conference
on Modeling and Using Context (pp. 11-23).

@@

We refer to \emph{Partner modelling} as the process of inferring one's partner
mental states. Any claim that students carry out a detailed monitoring of their
peers would be as incorrect as any claim that they do not maintain any
representation at all. If mutual modelling had to be permanently detailed and
accurate, subjects would obviously face a huge cognitive load. Conversely, peers
could not collaborate without some minimal amount of mutual modelling.
Collaborative learning dialogue include many instances of utterances such as ``I
thought he would do that'' (first level of mutual modelling) or even ``He
thought I would do that but I intended something else.'' (second level of mutual
modelling).

%For instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
%opinion. The mutual model can be \emph{implicit} ($A$ is not aware of what he knows about
%B), \emph{by-default} (I believe that $B$ believes what I believe unless contrary
%evidence), \emph{opportunistic} ($A$ does not model $B$ unless the conversation requires
%it), \emph{global} ($A$ infers $B$'s beliefs based on categories such as age, culture or
%profession) and, of course, it can be incorrect\ldots but it can not remain empty.

The content of the partner model ranges from \emph{dispositional} to
\emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
representation of $B$'s long term knowledge, skills or traits. It is thus closely
related to the notion of transactive memory~\citep{wegner1987transactive,
moreland1999transactive}.  \emph{Situational} aspects refer to $A$'s representation of
B's knowledge, behavior or intentions specifically activated in the situation in
which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
ones for 2 hours. Examples of fragments that constitute $A$'s model of B
regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:

@@

    \item $Model(A,B, knowledge)$: what does $A$ know about $B$'s knowledge with
        respect to the task at hand or, inversely, about $B$'s knowledge gaps?
        When can $A$ consider $B$'s statements as reliable?

    \item $Model(A,B, skills)$: what does $A$ know about $B$'s skills with respect to
        the task at hand? May $A$ expect $B$ to perform well in a specific subtask?
        The effectiveness of division of labor depends on the quality of this
        mutual model.

    \item $Model(A,B, goals)$: what does $A$ know about $B$'s intentions with respect
        to the project, including $B$' motivation and commitment? Can $A$ trust B
        when $B$ promises to deliver?

    \item $Model(A,B, task)$: what does $A$ know about $B$'s representation of the
        situation and the task: does $A$ knows whether $B$ has the same
        understanding of the problem at stake?

    \item $Model(A,B, plans)$: what does $A$ know about $B$'s strategy. Does A
        understand why $B$ did what he did? Is $A$ able to anticipate what $B$ will do
        next?

    \item $Model(A,B, ``urgent'')$: what does $A$ know about $B$'s understanding of $A$'s
        last utterance: does ``urgent'' means now, soon or ``not too late''?

@@

The list of what $X$ stands for in \model{A}{B}{X} is possibly infinite:
beliefs, emotions, history, status, etc. A partner model is likely not a
``box'', \ie not a monolithic representation but rather a mosaic of information
fragments about the partner, with various granularity and various life cycles.
This mosaic is elaborated through a variety of mechanisms, first for building an
initial model of the partner, then for updating this model. As two students meet
for the first time, partners models are initialized by the assumptions they make
upon each other based on cues such as his/her belonging to broad categories
(age, culture, profession, \ldots), stereotypes (sportsmen, junkie, business
women, Swiss,\ldots) as well as physical appearance. Scholars studied how
initial modelling impacts communication. In their experiment,
\citet{slugoski1993attribution} pretended to their subjects that their (fake)
partner had or had not received the same information. They observed that the
subjects adapted their dialogue by focusing the explanation on the items that
he/she was supposed to ignore.  \citet{brennan1991conversation} showed that the
subjects used different initial strategies in forming queries depending on who
they were told their partner was.

Initial common grounds are also initiated by co-presence: they include events to
which $A$ and $B$ attended together~\citep{clark2002definite} in the physical
space or in their cultural space (\eg ``09-11''). While co-presence means that
they can refer to shared objects and events, it does not imply that they give
them the same meaning. Namely, a shared screen does not mean a shared
understanding~\citep{dillenbourg2006sharing}.

@@

Mutual modelling is bi-directional. During dyadic problem solving,   partner A
builds some model of B and B build some model of A.  Moreover, these two
processes are not independent: A's model of what B knows, include what B knows
about A.  This leads to nested levels of  modelling. If $A$ states ``B thinks I
am good in maths'', $A$ builds a \emph{second level} model:
\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
regress of nested models: $A$ saying ``$B$ knows that I don't expect him to
solve this statistics problem'' corresponds to \\
\model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.  As we will see in
one study we report, mutuality also applies to triads in which A will elaborate
a model of B and of C, and reciprocally. This is probably also true for larger
groups although one may hypothesize that there exists (yet undefined) group size
from which it is not possible to model all partners individually and, therefore,
members  model  the group as a whole instead as a collection of individuals.  We
do not explore this hypothesis here and limit ourselves to dyads and triads.

This mutuality allows us to address a fundamental question: \emph{does the
quality of partner modelling depend upon the cognitive skills of each partner
(some people being better in perceiving other's states) or does it result from
the quality of interactions among them ?} Behind his question, the reader may
perceive the long lasting debate between tenets of, respectively, the individual
and social views of human cognition.  The simple hypothesis is that when
individual are randomly paired for an experiment, there is no reason for which
their individual cognitive skills would correlate. Therefore, if it occurs that
the quality of the Model (A,B) is correlated with Model (B,A), one may infers
that this quality depends upon what A and B have in common, their verbal
interactions.

@@

Natural language becomes cumbersome when describing things such as "the model
that A builds about the model that B builds about A". Therefore, to define
hypotheses and report on experiments on mutual modelling, we use  the notation
\model{A}{B}{X} to denote ``$A$ knows that $B$ knows X''. This  notation is not
proposed as a formal theory of mutual modelling but as useful simplification for
communicating about mutual modelling. This notation does not mean $A$ has an
explicit, monolithic representation of $B$: it must be understood as an
abstraction referring to complex socio-cognitive processes.  As explained in the
previous section, $A$ can be fragmented, multi-dimensional, etc. This notation
is neither presented as a computational model of mutual modelling, nor as some
universal formalism; it's usefulness is internal to this paper.  Besides, we
refer to the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We
discuss in the next section the methodological difficulty in measuring this
accuracy.

@@

    \item Before any specific grounding action, there is generally a non-null
        probability that $X$ is mutually understood by $A$ and $B$ (\eg $X$ is
        part of $A$'s and $B$'s cultures, it is manifest to co-present subjects
        or simply there is not much space for misunderstanding or disagreement
        about $X$). We simply could not collaborate without a certain level of
        initial grounds. We note the theoretical accuracy of \emph{initial
        grounds} $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.

@@

        \item \model{A}{B}{actions} is about how well $A$ guesses what action $B$ has
            performed (study 2) or will perform next (study 1),

        \item \model{A}{B}{emotion}: how accurately $A$ perceives $B$'s emotional state
            (study 3),

        \item \model{A}{B}{knowledge}: how accurately $A$ estimates $B$'s knowledge
            with respect to the material they learn together (study 4 and 5).

@@

    \item $cost(\alpha)$: in all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provides $A$ with information about
        what $B$ does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
        and, like any prosthesis, they may augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

@@

{\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$ what he/she
believes about what $B$ knows, feels, intends to do, etc.  This raises obvious
methodological concerns since such a question triggers a modelling process
beyond what would naturally occur. To avoid this bias, one can estimate mutual
modelling after task completion.  Then, the obvious drawback are memory losses
and by post-hoc reconstruction.  The first method was used in study 2 and the
second one in the other studies. Another option would be to rely on external
behavioural metrics like eye-tracking: we hypothesise for instance that the
fixation time reflects the efforts engaged by the human to understand, hence,
model, the others. Such an approach has however not been investigated in the
presented studies.

{\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we need to
access the reference model \refmodel{B}{X} to estimate its \emph{accuracy}.
Since we can only indirectly access it via what $B$ reports (\ie
\model{B}{B}{X}), accuracy can be estimated in 2 ways:

@@

    \item Subjective accuracy: In study 3, we compute \Model{A}{B}{X} by
        measuring if $A$ describes $B$'s emotions in the same way $B$ reports
        its emotions ($\M{A}{B}{X} = \M{B}{B}{X}$).

    \item Objective accuracy: In studies 4 and 5, we compute \Model{A}{B}{X} by
        comparing \Model{A}{B}{K} to $B$'s actual knowledge as it has measured
        by a test.

@@

Our method for investigating mutual modelling is observe the variations of
accuracy that result from variations of external parameters that appear as
variable in the formula presented in the previous section: the accuracy should
for go down if we increase the cost of grounding acts and goes up if we increase
the grounding criterion (the necessity to build a shared understanding). One way
to produce these variations is what CSCW researchers have called ``awareness
tools'', \ie functionalities that inform $A$ is about $B$'s actions that A could
not directly perceive because $B$ was working on a different subset of the
virtual space. Different awareness tools are used in the following studies as
methodological tricks  to manipulate experimentally the  mutual modelling
activity.

@@

As stated earlier, the fundamental challenge is to determine if mutual modelling
is an individual skill (hypothesis 1) or  the emergent property of social
interactions (hypothesis 3).  This question is empirically translated into the
symmetry of mutual modelling: what is the relationship between \Model{A}{B}{X}
and \Model{B}{A}{X} (Fig.~\ref{mm_symmetry})? A low symmetry would mean that
mutual modelling is mainly an individual attitude/aptitude ($\mathcal{H}_{1}$).
A high correlation might support $\mathcal{H}_{3}$ since there is a low
probability that randomly formed pairs integrate peers with the same level of
mutual modelling skills. However, a high correlation could also have another
explanation, stated in  $\mathcal{H}_{2}$: it could be that A is good at
modelling B and good at helping B to model herself or himself.

>>>>

\usepackage{xspace}
\newcommand{\ie}{i.e.\xspace}
\newcommand{\cf}{{\textit{cf\xspace}}}
\newcommand{\eg}{{\textit{e.g.\xspace}}}
\newcommand{\al}{{\textit{et al.\xspace}}}

\newcommand{\A}{A\xspace}
\newcommand{\B}{B\xspace}
\newcommand{\C}{C\xspace}

@@

Collaborative learning has often been associated to the construction of  a
shared   understanding of the situation at hand. The psycholinguistics
mechanisms at work while establishing common grounds are the object of
scientific controversy. We postulate that collaborative tasks require some
level of mutual modelling, \ie that each partner needs some model of what
the other partners know/want/intend at a given time.  We use the term ``some
model'' to stress the fact that this model is not necessarily detailed or
complete, but that we acquire some representations of the persons we
interact with. The question we address is: Does the quality of the partner
model depend upon the modeler's ability to represent his or her partner?
Upon the modelee's ability to make his state clear to the modeler? Or rather upon
the quality of their interactions? We address this question by comparing
the respective accuracies of the models built by different team
members. We therefore report on 5 experiments on collaborative problem solving
or collaborative learning that vary in terms of tasks (how important it is
to build an accurate model) and settings (how difficult it is to build an
accurate model). In 4 studies, the accuracy of the model that \A built about
\B was correlated with the accuracy of the model that \B built about \A,  which
seems to imply that the quality of interactions matters more than individual
abilities when building mutual models.  However, these finding do not
rule out the fact that individual abilities also participate in the quality of
modelling process.

@@

From its inception, Computer Supported Collaborative Learning (CSCL)
research has been following the suggestion by \citet{roschelle1995construction}
that collaborative learning has something to do with the process of constructing
and maintaining a \emph{shared understanding} of the task at hand. Building a
shared/mutual understanding refers to the upper class of collaborative learning
situations, those in which students should build upon each other's understanding
to refine their own understanding: what is expected to produce learning is not
the mere fact that two students build the same understanding but the cognitive
effort they have to engage to build this shared
understanding~\citep{schwartz1995emergence}. This effort can be observed by the
frequency of rich interactions, \ie interactions whose occurrence has been
related to learning: (self-) explanations in cognitive
science~\citep{chi1989self, webb1991task}, conflict resolution in
socio-cognitive theories~\citep{doise1975social}, mutual
regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
construction of a shared understanding has been investigated for several decades
in psycholinguistics, under the  notion of
\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned in the learning
sciences. Grounding mechanisms are appropriate to explain conversational
events, such as referential failures in short dialogue episodes, but they do hardly
predict deeper phenomena such as \emph{conceptual change} (\ie the acquisition,
acceptation and integration of a new belief into one's mental model) over longer
sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:

@@

Paradoxically, while Clark's theory is somewhat too linguistic from a learning
viewpoint, it is criticized at the same time as being too cognitivist by some
psycholinguists, \ie as overestimating the amount of shared knowledge and mutual
representations actually necessary to conduct a dialogue. The fundamental issue,
as old as philosophy, is the degree of coupling between the different levels of
dialogue, mostly between the lexical/syntactical level and the deeper semantic
levels. \citet{pickering2006alignment} argue that mutual understanding
starts mostly with a \emph{superficial alignment} at the level of the linguistic
representations, due to priming mechanisms, and that this local alignment may --
in some cases -- lead to a \emph{global alignment} of the semantic level
(\emph{deep grounding}).  For these authors, the convergence in dialogue, and
even the repair of some misunderstandings, is explained by this mimetic behavior
more than by a monitoring of each other's knowledge: \emph{``...interlocutors do
not need to monitor and develop full common ground as a regular, constant
part of routine conversation, as it would be unnecessary and far too costly.
Establishment of full common ground is, we argue, a specialized and
non-automatic process that is used primarily in times of difficulty (when
radical misalignment becomes apparent).''} (p. 179).  This view is actually not
incompatible with Clark's \emph{grounding
criterion}~\citep{clark1989contributing}: the degree of shared understanding
that peers need to reach depends upon the task they perform. For instance, a
dialogue between two surgeons might rely on superficial alignment if they talk
about their friends but has to guarantee accurate common grounds when talking
about which intervention will be conducted in which way on which patient.  In
this paper, we operationalized the grounding criterion, \ie the necessity for
accurate modelling, as \emph{the correlation between the accuracy of partner models
and measures of team performance}.

This interesting cognitive science debate occurred mostly outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes: Is linguistic alignment sufficient to trigger conceptual
change? Does negotiation of meaning only occur when partners monitor
and diagnose each other's knowledge? If the ratio between shallow alignment and
deep grounding depends upon the task, and if deep grounding is a condition for
learning, then the pedagogical challenge is indeed to design tasks that require
deep grounding. Most empirical studies on grounding and alignment are conducted
with simple referencing tasks such as asking the way to the train station
or helping the peer to choose a picture among many. In the studies we report here, we
explore several richer tasks such as arguing on a sensitive issue or building a
concept map.

@@

%\paragraph{Theory of Mind}
%
%Theory of Mind (originally defined in~\citep{premack1978does}) is the cognitive
%ability that a subject possesses to represent the mental state of another
%agent, possibly including knowledge that contradicts the subject's own model: for
%example, a book can be at the same time \emph{visible} for myself, and \emph{not
%visible} for you.
%
%Children develop this skill, which is essential to understand others'
%perspectives during interactions, around the age of three. It supposes the
%ability to build, store and retrieve separate models of the knowledge of the
%interactors.
%
%One classical application of this cognitive skill is the so-called
%\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
%experiment)~\citep{Leslie2000}: a child is asked to watch a scene where two
%people, \A and \B, manipulate objects. Then \A leaves and \B hides away one
%object. When \A comes back, we ask the child ``where do you think \A will
%look for the object?''. Before acquiring a theory of mind, children are not
%able to separate their own (true) model of the world (where they know that
%the object was hidden) from the model of A, which contains \emph{false
%beliefs} on the world (\A still thinks the object is at its original
%position since he did not see \B hiding it).

@@

In summary, the controversy around the cognitive depth of shared understanding
pertains to psycholinguistic, which investigates natural conversations. The
situation is different in collaborative learning; tasks are actually designed
for requiring a deeper negotiation of meaning.  The expertise of CSCL is to design
collaborative situations that create interdependence, avoid group think, and
allows students to detect any ``illusion of shared
understanding''~\citet{cherubini2005grounding}. Our question is hence not
anymore ``do peers build a shared
understanding ?'', but rather ``whenever peers have to build a shared understanding,
how do they achieve it ?''.  This question addresses the mechanisms of grounding, in
which the basic sequence is: make a proposition, detect misunderstanding and, if
any, repair them.  This paper focuses on the middle part, the detection of
misunderstandings.  Detecting a misunderstanding means that the emitter of a
message identifies a mismatch between his communicative intention and the way
his message is understood by his or her partner. Detecting one peer's
misunderstanding is investigated hereafter under the general umbrella of partner
modelling.

@@

We refer to \emph{Partner modelling} as the process of inferring one's partner
mental states. Any claim that students carry out a detailed monitoring of their
peers would be as incorrect as any claim that they do not maintain any
representation at all. If mental modelling had to be permanently detailed and
accurate, subjects would obviously face a huge cognitive load. Conversely, peers
could not collaborate without some minimal amount of mutual modelling.
Collaborative learning dialogue include many instances of utterances such as ``I
thought he would do that'' (first order modelling) or even ``He
thought I would do that but I intended something else.'' (second order
modelling).

%For instance, \A cannot disagree with \B without knowing that \B has a different
%opinion. The mutual model can be \emph{implicit} (\A is not aware of what he knows about
%B), \emph{by-default} (I believe that \B believes what I believe unless contrary
%evidence), \emph{opportunistic} (\A does not model \B unless the conversation requires
%it), \emph{global} (\A infers \B's beliefs based on categories such as age, culture or
%profession) and, of course, it can be incorrect\ldots but it can not remain empty.

The content of the partner model ranges from \emph{dispositional} to
\emph{situational} aspects. The \emph{dispositional} aspects refer to \A's
representation of \B's long term knowledge, skills or traits. It is thus closely
related to the notion of transactive memory~\citep{wegner1987transactive,
moreland1999transactive}.  \emph{Situational} aspects refer to \A's representation of
B's knowledge, behavior or intentions specifically activated in the situation in
which \A and \B are collaborating, some of them being valid for 2 seconds, other
ones for 2 hours. Examples of fragments that constitute \A's model of B
regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:

@@

    \item $Model(A,B, knowledge)$: what does \A know about \B's knowledge with
        respect to the task at hand or, inversely, about \B's knowledge gaps?
        When can \A consider \B's statements as reliable?

    \item $Model(A,B, skills)$: what does \A know about \B's skills with respect to
        the task at hand? May \A expect \B to perform well in a specific subtask?
        The effectiveness of division of labor depends on the quality of this
        mutual model.

    \item $Model(A,B, goals)$: what does \A know about \B's intentions with respect
        to the project, including \B' motivation and commitment? Can \A trust
        \B
        when \B promises to deliver?

    \item $Model(A,B, task)$: what does \A know about \B's representation of the
        situation and the task: does \A knows whether \B has the same
        understanding of the problem at stake?

    \item $Model(A,B, plans)$: what does \A know about \B's strategy. Does A
        understand why \B did what he did? Is \A able to anticipate what \B will do
        next?

    \item $Model(A,B, ``urgent")$: what does \A know about \B's understanding of \A's
        last utterance: does ``urgent'' means now, soon or ``not too late''?

@@

The list of what $X$ stands for in \model{A}{B}{X} is possibly infinite:
beliefs, emotions, history, status, etc. A partner model is likely not a
``box'', \ie not a monolithic representation but rather a mosaic of information
fragments about the partner, with various granularity and various life cycles.
This mosaic is elaborated through a variety of mechanisms, first for building an
initial model of the partner, then for updating this model. As two students meet
for the first time, partners models are initialized by the assumptions they make
upon each other based on cues such as his/her belonging to broad categories
(age, culture, profession, \ldots), stereotypes (sportsmen, junkie, business
women, Swiss,\ldots) as well as physical appearance. Scholars studied how
initial modelling impacts communication. In their experiment,
\citet{slugoski1993attribution} pretended to their subjects that their
(confederate) partner had or had not received the same information. They
observed that the subjects adapted their dialogue by focusing the explanation on
the items that he/she was supposed to ignore.  \citet{brennan1991conversation}
showed that the subjects used different initial strategies in forming queries
depending on who they were told their partner was.

Initial common grounds are also initiated by co-presence: they include events to
which \A and \B attended together~\citep{clark2002definite} in the physical
space or in their cultural space (\eg ``09-11''). While co-presence means that
they can refer to shared objects and events, it does not imply that they give
them the same meaning. Namely, a shared screen does not mean a shared
understanding~\citep{dillenbourg2006sharing}.

@@

Mutual modelling is bi-directional. During dyadic problem solving,   partner A
builds some model of \B and \B build some model of \A.  Moreover, these two
processes are not independent: \A's model of what \B knows, include what \B knows
about \A.  This leads to nested levels of  modelling. If \A states ``\B thinks I
am good in maths'', \A builds a \emph{second level} model:
\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
regress of nested models: \A saying ``\B knows that I don't expect him to
solve this statistics problem'' corresponds to \\
\model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.  As we will see in
one study we report on, mutuality also applies to triads in which \A will elaborate
a model of \B and of C, and reciprocally. This is probably also true for larger
groups although one may hypothesize that there exists (yet undefined) group size
from which it is not possible to model all partners individually and, therefore,
members  model  the group as a whole instead as a collection of individuals.  We
do not explore this hypothesis here and limit ourselves to dyads and triads.

This mutuality allows us to address a fundamental question: \emph{does the
quality of partner modelling depend upon the cognitive skills of each partner
(some people being better in perceiving other's states) or does it result from
the quality of interactions among them ?} Behind his question, the reader may
perceive the long lasting debate between tenets of, respectively, the individual
and social views of human cognition.  The simple hypothesis is that when
individual are randomly paired for an experiment, there is no reason for which
their individual cognitive skills would correlate. Therefore, if it occurs that
the quality of the $Model(A,B)$ is correlated with $Model(B,A)$, one may infers
that this quality depends upon what \A and \B have built together while
interacting.

@@

Natural language becomes cumbersome when describing things such as ``the model
that \A builds about the model that \B builds about A''. Therefore, to define
hypotheses and report on experiments on mutual modelling, we use  the notation
\model{A}{B}{X} to denote ``\A knows that \B knows X''. This  notation is not
proposed as a formal theory of mutual modelling but as useful simplification for
communicating about mutual modelling. This notation does not mean \A has an
explicit, monolithic representation of \B: it must be understood as an
abstraction referring to complex socio-cognitive processes.  As explained in the
previous section, \A can be fragmented, multi-dimensional, etc. This notation
is neither presented as a computational model of mutual modelling, nor as some
universal formalism; it's usefulness is internal to this paper.  Besides, we
refer to the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We
discuss in the next section the methodological difficulty in measuring this
accuracy.

@@

    \item Before any specific grounding action, there is generally a non-null
        probability that $X$ is mutually understood by \A and \B (\eg $X$ is
        part of \A's and \B's cultures, it is manifest to co-present subjects
        or simply there is not much space for misunderstanding or disagreement
        about $X$). We simply could not collaborate without a certain level of
        initial grounds. We note the theoretical accuracy of \emph{initial
        grounds} $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.

@@

        \item \model{A}{B}{actions} is about how well \A guesses what action \B has
            performed (study 2) or will perform next (study 1),

        \item \model{A}{B}{emotion}: how accurately \A perceives \B's emotional state
            (study 3),

        \item \model{A}{B}{knowledge}: how accurately \A estimates \B's knowledge
            with respect to the material they learn together (study 4 and 5).

@@

    \item $cost(\alpha)$: in all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provides \A with information about
        what \B does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
        and, like any prosthesis, they may augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

@@

{\bf Capturing \model{A}{B}{X}}: The simplest method is to ask \A what he/she
believes about what \B knows, feels, intends to do, etc.  This raises obvious
methodological concerns since such a question triggers a modelling process
beyond what would naturally occur. To avoid this bias, one can estimate mutual
modelling after task completion.  Then, the obvious drawback are memory losses
and by post-hoc reconstruction.  The first method was used in study 2 and the
second one in the other studies. Another option would be to rely on external
behavioural metrics like eye-tracking: we hypothesise for instance that the
fixation time reflects the efforts engaged by the human to understand, hence,
model, the others. Such an approach has however not been investigated in the
presented studies.

{\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we need to
access the reference model \refmodel{B}{X} to estimate its \emph{accuracy}.
Since we can only indirectly access it via what \B reports (\ie
\model{B}{B}{X}), accuracy can be estimated in 2 ways:

@@

    \item Subjective accuracy: In study 3, we compute \Model{A}{B}{X} by
        measuring if \A describes \B's emotions in the same way \B reports
        its emotions ($\M{A}{B}{X} = \M{B}{B}{X}$).

    \item Objective accuracy: In studies 4 and 5, we compute \Model{A}{B}{X} by
        comparing \Model{A}{B}{K} to \B's actual knowledge as it has measured
        by a test.

@@

Our method for investigating mutual modelling is observe the variations of
accuracy that result from variations of external parameters that appear as
variable in the formula presented in the previous section: the accuracy should
for go down if we increase the cost of grounding acts and goes up if we increase
the grounding criterion (the necessity to build a shared understanding). One way
to produce these variations is what CSCW researchers have called ``awareness
tools'', \ie functionalities that inform \A is about \B's actions that \A could
not directly perceive because \B was working on a different subset of the
virtual space. Different awareness tools are used in the following studies as
methodological tricks  to manipulate experimentally the  mutual modelling
activity.

@@

As stated earlier, the fundamental challenge is to determine if mutual modelling
is an individual skill (hypothesis 1) or  the emergent property of social
interactions (hypothesis 3).  This question is empirically translated into the
symmetry of mutual modelling: what is the relationship between \Model{A}{B}{X}
and \Model{B}{A}{X} (Fig.~\ref{mm_symmetry})? A low symmetry would mean that
mutual modelling is mainly an individual attitude/aptitude ($\mathcal{H}_{1}$).
A high correlation might support $\mathcal{H}_{3}$ since there is a low
probability that randomly formed pairs integrate peers with the same level of
mutual modelling skills. However, a high correlation could also have another
explanation, stated in  $\mathcal{H}_{2}$: it could be that \A is good at
modelling \B and good at helping \B to model herself or himself.

\begin{itemize}
    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon \A's ability or effort
        to model \B,
